SeedLing: Building and using a seed corpus \\ for the Human Language Project

### 2014-04-28 docs/computEL-final/clean.tex
Further edits

\paragraph{UDHR.} We used the plaintext UDHR files available from the Unicode website\footnote{\url{http://unicode.org/udhr/d}} which uses UTF-8 encoding for all languages. The first four lines of each file record metadata, and the rest is the translation of the UDHR. This dataset is extremely clean, and simply required segmentation into sentences.

@@

\subsection{Language Codes}
As \newcite{xia2010multilingual} explain, language names do not always suffice to identify languages, since many names are ambiguous. For this reason, sets of language codes exist to more accurately identify languages. We use ISO~639-3\footnote{url{http://www-01.sil.org/iso639-3/default.asp}} as our standard set of codes, since it aims for universal coverage, and has widespread acceptance in the community. The data from ODIN and the UDHR already used this standard.

>>>>

\paragraph{UDHR.} We used the plain-text UDHR files available from the Unicode website\footnote{\url{http://unicode.org/udhr/d}} which uses UTF-8 encoding for all languages. The first four lines of each file record metadata, and the rest is the translation of the UDHR. This dataset is extremely clean, and simply required segmentation into sentences.

@@

\subsection{Language Codes}
As \newcite{xia2010multilingual} explain, language names do not always suffice to identify languages, since many names are ambiguous. For this reason, sets of language codes exist to more accurately identify languages. We use ISO~639-3\footnote{\url{http://www-01.sil.org/iso639-3/default.asp}} as our standard set of codes, since it aims for universal coverage, and has widespread acceptance in the community. The data from ODIN and the UDHR already used this standard.

### 2014-04-28 docs/computEL-final/cluster.tex
Further edits

\paragraph{Experimental Setup.}
We performed hierarchical clustering, which produces a tree structure: each leaf represents a language, and each node a cluster. We used linkage methods, which recursively build the tree starting from the leaves. Initially, each language is in a separate cluster, then we iteratively find the closest two clusters and merge them. Each time we do this, we take the two corresponding subtrees, and introduce a new node to join them.

We tested four methods, which differ in how they define the distance between two clusters. The \texttt{complete} method considers all pairs of languages with one from each cluster, and finds the largest distance. The \texttt{single} method instead finds the smallest distance. The \texttt{average} and \texttt{weighted} methods calculate averages across all the pairwise differences, either unweighted, or weighted according to the languages' positions in the tree.

While a tree structure might be linguistically useful, evaluating the structure is more awkward. However, the above methods can be easily used to produce partitional clusters, by stopping when we reach a certain number of clusters. We set the number of clusters to 147, the number of top-level genetic groupings in Ethnologue.

@@

    \begin{tabular}{l|ccc}
    ~        & Precision & Recall       & F-score    \\ \hline
	\texttt{complete} & \textbf{0.3142}	& 0.1938	 & \textbf{0.1556} \\
	\texttt{single} & 0.1949	& \textbf{0.7850}	 & 0.1282 \\
	\texttt{average} & 0.2216	& 0.5163	& 0.1348 \\
	\texttt{weighted} & 0.2326	& 0.3343	& 0.1271 \\ \hline
	\emph{random} & 0.1836	& 0.0917	& 0.0675 \\
    \end{tabular}
\caption{Comparison of clustering algorithms}
\label{table:cluster}
\end{centering}
\end{table}

\paragraph{Evaluation.}
There are many possible metrics to evaluate the quality of a clustering compared to a gold standard. 
\newcite{amigo2009metrics} propose a set of criteria which a clustering evaluation metric should satisfy, and demonstrate that most popular metrics fail to satisfy at least one of these criteria.  However, they prove that they are satisfied by the BCubed metric, which we adopt for this reason.  To calculate this, we take the induced cluster and gold standard class for each language, and calculate the F-score of the cluster compared to the class.  These F-scores are then averaged across all languages.

@@

To allow a closer comparison with \newcite{georgi2010wals}, we can also calculate pairwise scores - i.e. considering when pairs of languages are in the same cluster or the same class. For 147 clusters, the \texttt{complete} method gives a pairwise f-score of 0.147, while they report 0.140. The figures are not quite comparable since we are evaluating over a different set of languages; nonetheless, this result is suggestive, given that we have only used surface features, while they used typological features taken from WALS.  This demonstrates that it possible for cross-linguistic research to be conducted even based on extremely shallow features.

@@

\paragraph{Qualitative Inspection.}
Ga: It's actually pretty difficult to interpret the clusters directly because there's so much noise. I will try to write something better, but here are a few thoughts I've had:

Using the \texttt{complete} method with 147 clusters, we obtain a mix of large and small clusters. There are 27 singleton clusters, while the largest cluster contains 123 languages. Although these singleton clusters belong to classes of various size, it is interesting that the distribution of sizes is similar to the distribution in the gold standard data, where we have 27 single classes, and the largest class contains 150 languages.  This isn't true for the other clustering methods - the \texttt{single} method produces 138 singleton clusters, with almost all languages in a single cluster of size 968.

>>>>

\paragraph{Experimental Setup.}
We first perform hierarchical clustering, which produces a tree structure: each leaf represents a language, and each node a cluster. We use linkage methods, which recursively build the tree starting from the leaves. Initially, each language is in a separate cluster, then we iteratively find the closest two clusters and merge them. Each time we do this, we take the two corresponding subtrees, and introduce a new node to join them.

We test four methods, which differ in how they define the distance between two clusters. The \texttt{complete} method considers all possible pairs with one language from each cluster, and finds the largest distance. The \texttt{single} method instead finds the smallest distance. The \texttt{average} and \texttt{weighted} methods calculate averages across all the pairwise differences, either unweighted, or weighted according to the languages' positions in the tree.

To ease evaluation, we produce a partitional clustering, by stopping when we reach a certain number of clusters. We set the number of clusters to 147, the number of top-level genetic groupings in Ethnologue.

@@

    \begin{tabular}{l|ccc}
    ~        & Precision & Recall       & F-score    \\ \hline
	\emph{random} & 0.1836	& 0.0917	& 0.0675 \\ \hline
	\texttt{complete} & \textbf{0.3142}	& 0.1938	 & \textbf{0.1556} \\
	\texttt{single} & 0.1949	& \textbf{0.7850}	 & 0.1282 \\
	\texttt{average} & 0.2216	& 0.5163	& 0.1348 \\
	\texttt{weighted} & 0.2326	& 0.3343	& 0.1271 \\
    \end{tabular}
\caption{Comparison of clustering algorithms}
\label{table:cluster}
\end{centering}
\end{table}

\paragraph{Evaluation.}
We compare our induced clusters to the language families given in Ethnologue. However, there are many possible metrics to evaluate the quality of a clustering. \newcite{amigo2009metrics} propose a set of criteria which a clustering evaluation metric should satisfy, and demonstrate that most popular metrics fail to satisfy at least one of these criteria.  However, they prove that all criteria are satisfied by the BCubed metric, which we adopt for this reason.  To calculate the BCubed score, we take the induced cluster and gold standard class for each language, and calculate the F-score of the cluster compared to the class.  These F-scores are then averaged across all languages.

@@

To allow a closer comparison with \newcite{georgi2010wals}, we can also calculate pairwise scores - i.e. considering when pairs of languages are in the same cluster or the same class. For 147 clusters, the \texttt{complete} method gives a pairwise f-score of 0.147, while Georgi et al. report 0.140. The figures are not quite comparable since we are evaluating over a different set of languages; nonetheless, this result is suggestive, given that we only use surface features, while Georgi et al. use typological features taken from WALS.  This demonstrates that it possible for cross-linguistic research to be conducted even based on extremely shallow features.

@@

\paragraph{Qualitative Inspection.}
Interpreting the clusters directly is difficult, because of a large amount of noise in the data.
Using the \texttt{complete} method with 147 clusters, we obtain a mix of large and small clusters. There are 27 singleton clusters, while the largest cluster contains 123 languages. Although these singleton clusters belong to classes of various size, it is interesting that the distribution of sizes is similar to the distribution in the gold standard data, where we have 27 single classes, and the largest class contains 150 languages.  This isn't true for the other clustering methods - the \texttt{single} method produces 138 singleton clusters, with almost all languages in a single cluster of size 968.

### 2014-04-28 docs/computEL-final/data.tex
Further edits

Building on their previous paper, \newcite{abney2011data} describe the data structure they envisage for the universal corpus in more detail, distinguishing between \textbf{aligned texts} and \textbf{analysed texts}. Aligned texts consist of multiple parallel documents, aligned at the document, sentence, or word level. The collection of parallel documents as a whole is assigned a unique identifier, with each individual document labelled by this identifier and the corresponding language code.  A monolingual document is simply one with only a single associated language code.  Sentences and words within documents are also assigned identifiers (relative to the document), if alignment has been performed at that level.

Analysed texts, in addition to the raw text, contain more detailed annotations including parts of speech, morphological information, and syntactic relations.  This is stored as a table, where each row represents a word, and columns can represent the following types of information: document ID, language code, sentence ID, word ID, wordform, lemma, morphological information, part of speech, gloss, head/governor, and relation/role.

@@

\begin{table}[t]
\small
\centering
    \begin{tabular}{l|rr|rr}
    ~         				& \#Languages & \#Families 	&\#tokens		& Size	\\ \hline
    ODIN      				& 1,217      & 101       		& 58,556		& 39 MB		\\
    Omniglot  				& 134        & 21        		&	6,749			& 677 KB	\\
    UDHR      				& 355        & 47        		&	33,117		& 5.2 MB	\\
    Wikipedia 				& 209        & 22       		&						& 37 GB		\\ \hline
    \textbf{Combined}	& 1,347			 & 106 
    \end{tabular}
\caption{Corpus Coverage}
\label{table:corpus}
\end{table}

>>>>

Building on their previous paper, \newcite{abney2011data} describe the data structure they envisage for the universal corpus in more detail, and we aim to adopt this structure where possible.  Two types of text are distinguished: \textbf{aligned texts} consist of multiple parallel documents, aligned at the document, sentence, or word level. The collection of parallel documents as a whole is assigned a unique identifier, with each individual document labelled by this identifier and the corresponding language code.  A monolingual document is simply one with only a single associated language code.  Sentences and words within documents are also assigned identifiers (relative to the document), if alignment has been performed at that level.

Analysed texts, in addition to the raw text, contain more detailed annotations including parts of speech, morphological information, and syntactic relations.  This is stored as a table, where rows represent words, and columns represent the following types of information: document ID, language code, sentence ID, word ID, wordform, lemma, morphological information, part of speech, gloss, head/governor, and relation/role.

@@

\begin{table}[t]
\small
\centering
    \begin{tabular}{l|rr|rr}
    ~         				& Langs. 	& Families 	& Tokens		& Size	\\ \hline
    ODIN      				& 1,217      & 101       		& 58,556		& 39 MB		\\
    Omniglot  				& 134        & 21        		&	6,749			& 677 KB	\\
    UDHR      				& 355        & 47        		&	33,117		& 5.2 MB	\\
    Wikipedia 				& 209        & 22       		&						& 37 GB		\\ \hline
    \textbf{Combined}	& 1,347			 & 106 
    \end{tabular}
\caption{Corpus Coverage}
\label{table:corpus}
\end{table}

### 2014-04-28 docs/computEL-final/paper.tex
Further edits

\author{Guy Emerson, Liling Tan, Susanne Fertmann, Alexis Palmer and Michaela Regneri \\
Universität des Saarlandes \\
Campus, 66123 Saarbrücken, Germany \\
{\tt liling.tan@uni-saarland.de, emerson@coli.uni-saarland.de,} \\
{\tt susfert@coli.uni-saarland.de, apalmer@coli.uni-sb.de,}\\
{\tt regneri@coli.uni-saarland.de}}

\date{}

>>>>

\author{Guy Emerson, Liling Tan, Susanne Fertmann, Alexis Palmer and Michaela Regneri \\
Universität des Saarlandes \\
66123 Saarbrücken, Germany \\
{\tt \{emerson, liling, susfert, apalmer, regneri\}} \\
{\tt @coli.uni-saarland.de} \\
}
\date{}

### 2014-04-28 docs/computEL-final/relwork.tex
Further edits

In this section, we review existing efforts to compile multilingual machine-readable resources.  Although some commercial resources are available, we restrict attention to freely accessible data\footnote{All figures given below were correct at the time of writing, but it must be borne in mind that most resources discussed below are constantly growing.}.

@@

Such resources are invaluable for many purposes. For instance, there can be plenty of material to work with for a
researcher who is looking to find data on a specific language, who is willing to put in a little time browsing through different catalogues, and who only needs human-readable data.  However, for large-scale machine processing, they leave much to be desired.

@@

The Leipzig Corpora Collection (LCC)\footnote{\url{http://corpora.uni-leipzig.de}} \cite{biemann2007leipzig} provides download of corpora in 117 languages, and dictionaries in a number of others, bringing the total number of represented languages up to 230. The corpora are large, readily available, in plain text, and labelled with ISO language codes.

>>>>

In this section, we review existing efforts to compile multilingual machine-readable resources.  Although some commercial resources are available, we restrict attention to freely accessible data.\footnote{All figures given below were correct at the time of writing, but it must be borne in mind that most resources discussed below are constantly growing.}

@@

In short, while traditional archives are invaluable for many purposes, for large-scale machine processing, they leave much to be desired.

@@

The Leipzig Corpora Collection (LCC)\footnote{\url{http://corpora.uni-leipzig.de}} \cite{biemann2007leipzig} provides download of corpora in 117 languages, and dictionaries in a number of others, bringing the total number of represented languages up to 230. The corpora are large, readily available, in plain-text, and labelled with ISO language codes.
